Cultural Evolution of Conservation Behavior at the Intersection of Working

### 2020-12-15 DissertationProposal/DissertationProposal.Rmd
changed ch2 intro. Next is methods.

After the agents harvest resources from the landscape, the resources in the working and conservation areas will reproduce by the specified regeneration factor. If the number of resources in either area exceeds the allotted carrying capacity, the excess resources will either die (i.e. not be available for agents in the following time step) or move to the other area (conservation or working) with some specified probability also applied to a binomial distribution. Resources which move to the adjacent area but exceed the carrying capacity will also die. If we set the probability of resource mobility to 100%, we can now calculate the maximum sustainable harvest amount for each agent where the only modification to the equation is the addition of mobile resources from the conservation area (${P_{mob}}$) up until a specified carrying capacity ($W_{cc}$).

$$P_{mob} = (R*P_{r})-P_{r}$$

$$H_{sust} \leq \frac{(1-R)*(W_{r}-H_{max}*N)+min\{P_{mob},W_{cc}-W_{r}\}}{N} $$

@@

```{r abm2, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) for many combinationa of maximum harvest values and conservation area sizes. Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources. The regeneration factor at each time step is 1.20.", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/abmp2.png")
```

@@

## Background
Cultural group selection offers specific predictions about when individually costly, yet group beneficial norms will be favored by selection [@molleman_cultural_2013]. One such prediction is that the presence of multiple competing groups is the key to the development of norms for sustainable resource use [@waring_coevolution_2017]. Under this prediction, selection acts both on individuals and on groups of individuals. When the ratio of group benefit to individual cost of cooperation outweighs the individual fitness consequences of cooperation, those behaviors will be favored by selection. Groups who develop cooperative norms around natural resource use will outcompete other groups, who may then copy the cooperative institutions established by the more successful groups [@boyd_transmission_2010; @henrich_cultural_2004].

While these predictions are theoretically sound, they must be tested empirically [@kline2018designing]. Two such studies have begun to test the basis of cultural group selection in the natural world. The first demonstrated that culture offers greater fodder for group selection than do genetics. Researchers estimated the total variance in both cultural and genetic traits within and between groups of individuals. The proportion of variance in an allele or trait found between groups is referred to as genetic $F_{ST}$ or cultural $F_{ST}$ respectively. Findings indicate that between human groups, there is greater differentiation in culture than genes, indicating that group selection will act more readily on cultural traits [@bell_culture_2009].

A second empirical study tested the prediction that large scale human cooperation is in fact a product of competition between cultural groups. Researchers created pair-wise estimates of cultural $F_{ST}$ between 12 ethnic groups. These estimates of cultural differentiation were combined with a survey of individual willingness to cooperate with members of the other groups. Results of this study showed a strong association between cultural similarity and willingness to cooperate. This research indicates that the presence of variable cultural groups influences within-group cooperative behavior [@handley_human_2020].

@@

This project aims to extend the empirical study of cultural group selection to cooperation around environmental problems explicitly. Insights from cultural group selection may provide solutions to pressing environmental challenges of collective action. If validated experimentally, principles of cultural group selection may allow conservation scientists to better predict the viability of conservation initiatives.

Protected areas are an ideal study system for testing the mechanisms of cultural group selection in real world conservation. When a community adopts a protected area, like the Community Forest Management Agreements observed on Pemba Island, Tanzania, individuals agree to forego a portion of their personal harvest potential for the benefit of the community. Under cultural group selection, we only expect this behavior to emerge when the proportion of dissimilar to similar cultural traits between a group and its neighbors outweighs the benefit/cost ratio of engaging in cooperative conservation. Plainly put, communities should be more likely to adopt CoFMAs if they are culturally dissimilar to their surrounding communities. This project will test this theory by evaluating the perceived individual costs and group benefits of CoFMAs, within and between group cultural similarities (cultural $F_{st}$), and actual adoption of CoFMAs in forest dependent wards on Pemba Island [@kline2018designing].

### Question: 
Do Pemban wards with stronger cultural differentiation from surrounding wards adopt CoFMAs at a higher rate than wards who are more culturally similar to their neighbors?

### Predictions
* P1: Cultural differentiation of Pemban wards will predict CoFMA adoption only when the perceived individual costs of CoFMAs are relatively low.

* P2: High levels of forest dependence and perceived costs of CoFMAs will prevent CoFMA adoption, even when groups are considerably culturally dissimilar to adjacent groups.

>>>>

After the agents harvest resources from the landscape, the resources in the working and conservation areas will reproduce by the specified regeneration factor. If the number of resources in either area exceeds the allotted carrying capacity, the excess resources will either die (i.e. not be available for agents in the following time step) or move to the other area (conservation or working) with some specified probability also applied to a binomial distribution. Resources which move to the adjacent area but exceed the carrying capacity will also die. If we set the probability of resource mobility to 100%, we can now calculate the maximum sustainable harvest amount for each agent where the only modification to the equation is the addition of mobile resources from the conservation area (${C_{mob}}$) up until a specified carrying capacity ($W_{cc}$).

$$C_{mob} = (R*C_{r})-C_{r}$$

$$H_{sust} \leq \frac{(1-R)*(W_{r}-H_{max}*N)+min\{C_{mob},W_{cc}-W_{r}\}}{N} $$

@@

```{r abm2, out.width = "75%",echo=FALSE, fig.cap="Average harvest values accross five model runs for all agents in the final time step (30) for many combinationa of maximum harvest values and conservation area sizes. Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources. The regeneration factor at each time step is 1.20.", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/abmp2.png")
```

@@

## Background
Human social structures developed and still exist in nested groups, ranging from a nuclear family unit to a national or even multi-national identity. Group behavioral norms can have profound impacts on the wellbeing of group members and their environment, far beyond that of a single individual. Cultural group selection theorizes, and to a small extent has empirically demonstrated, the mechanisms which establish and perpetuate group-level norms [@henrich_cultural_2004; @bell_culture_2009]. Most relevant to community conservation is the development of norms which are beneficial for the group, but cause an individual to incur some personal cost; we refer to these as costly cooperative behaviors [@molleman_cultural_2013]. Central to theories around the adoption of costly cooperative behaviors is the insight that selection acts on both individuals and groups. Individuals may engage in personally costly behavior if it is beneficial for their group, which they also depend on for their wellbeing [@handley_human_2020]. The degree to which an individual depends on each of their nested groups is directly proportional to the personal costs they will incur for that group [@okasha_multilevel_2004; @price1972extension]. To the extent that conservation is a collaborative dilemma, these insights from cultural group selection may be paramount to implementing effective community based conservation [@waring2015multilevel; @kline2018designing].

Conservation is a particularly wicked collection action problem because the group level costs and benefits operate at a much longer time scale than those of the individual. The group level benefits of conservation are often much less certain than the immediate individual costs of forgoing resources. Humans tend to steeply discount the benefits of their actions as those benefits move further away in time and the benefits become more uncertain [@dasgupta2005uncertainty]. This hyperbolic discounting makes conservation norms very unlikely to emerge and persist under our current understanding of cultural group selection — the individual costs are high and group benefits are low. The role of institutions in human society is to reduce these barriers and to encourage collaborative behavior [@north1990institutions; @ostrom1990governing]. One particular institution aimed at promoting costly cooperative behaviors around conservation is REDD+ [@angelsen2017redd; @angelsen2008moving].

REDD+ in Zanzibar was designed to give performance based payments to communities who successfully reduce deforestation [@burgess2010getting]. In theory, these immediate, promised payments should alter the ratio of individual costs to group benefits so that costly cooperation is favored under cultural group selection. While this project was implemented in only a select few communities (Figure \@ref(fig:moon)), this setup should also promote the spread of community conservation norms to adjacent communities through success-biased imitation [@boyd_transmission_2010].

The REDD+ program on Pemba helped establish CoFMAs in 18 wards on the island, some of which have been successful in slowing and even stopping deforestation (Figure \@ref(fig:ForestCoverMap). The payments for the conservation efforts however were never given to communities. While this is ethically troublesome, it is interesting in light of the fact that ten additional wards on the island have subsequently elected to establish CoFMAs. This leaves open the question of why community conservation areas are spreading on the island while they are not expected to do so under cultural group selection. One explanation posited by Andrews and Borgerhoff Mulder [-@andrews_cultural_2018] actually hinges on a major critique of REDD+; leakage [@andam2008measuring].

@@

Leakage [as defined by @andam2008measuring] is a form of land use spillover where an intervention, such as the CoFMAs established on Pemba by REDD+, has a negative, counteracting effect on the targeted outcome variable in another place. On Pemba, we have anecdotal testimonies that individuals in wards with CoFMAs have simply shifted their harvest pressure onto adjacent wards. While dubious for the efficacy of REDD+ writ large, on Pemba, leakage may actually be driving the spread of community based conservation observed in the ten new communities which elected to create CoFMAs post REDD+ [@andrewsdoes]. Just as the performance based payments from REDD+ were designed to alter the individual/community cost/benefit matrix so that selection favors costly collaboration via immediate community payments, we hypothesize that leakage from wards with CoFMAs may also act to favor costly cooperative behavior as individuals must act to defend their access to resources they would not have otherwise.

If the spread of CoFMAs on Pemba is indeed a counterintuitive result of leakage, it may have considerable implications for REDD+ overall and community conservation areas broadly. To answer this, we first need to confirm that leakage is indeed occurring on Pemba as a result of individuals adhering to CoFMAs. Secondly, we need to evaluate the motivations of individuals who elected to establish CoFMAs in the ten wards who subsequently elected to do so. An alternative hypothesis is that individuals are copying the behaviors of more successful groups and are ambivalent towards or even unaware of leakage. The overall goal of this project is to quantify the degree to which leakage is driving the cascade of community conservation area adoption on Pemba.

### Question: 
Is leakage occurring as a result of individuals adhering to CoFMA regulations? If leakage is occurring, is it driving the spread of CoFMAs on Pemba?

### Predictions
* P1: Regardless of the demarcated size of the conservation area in each CoFMA, the leakage pressure on adjacent wards will be associated with the rate of deforestation, i.e. the extent to which individuals adhere to conservation area regulations.

* P2: In wards that have subsequently applied for CoFMAS, the primary motivator will be to combat leakage from adjacent wards, more so than copying their success or intrinsic motivations.
